Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants \thanks{Supported by EPSRC grant EP/J012564/1 and ERC project 280053 (CPROVER).}

### 2017-02-02 hscc2017/paper.tex
reviewed 4.1 Description of the Benchmarks
reviewed 4.3 Results

The first set of benchmarks uses the discrete model $G_1(z)$ of a cruise control
system for a car, and accounts for rolling friction, aerodynamic drag, and
the gravitational disturbance force~\cite{Astrom08}. 
%The plant model is given as follows:
%
%\begin{equation}
%\label{cruise-control-c1}
%G_1(z)=\frac{0.0264}{z-0.998} \nonumber
%\end{equation} 
%
The second set of benchmarks considers the discrete model $G_2(z)$ 
of a simple spring-mass damper plant~\cite{DBLP:conf/hybrid/WangGRJF16}. 
%where the discrete 
%plant dynamics is represented by the following $z$-expression:
%
%\begin{equation}
%\label{spring-mass-damper-g}
%G_2(z)=\frac{5\times{10^{-5}}z + 5\times{10^{-5}}}{z^2 - 2z + 1.0001}. \nonumber
%\end{equation} 
%
A third set of benchmarks uses the discrete model $G_{3}(z)$ for satellite attitude 
dynamics~\cite{Franklin15}, which require attitude (pose) control
for orientation of antennas and sensors w.r.t.~earth.  
%The satellite
%attitude control is typically used for three-axis attitude tracking, but
%here we consider only one axis at a time.  
%The motion equation for the
%one-axis system, disregarding disturbance torques, is given by
%
%\begin{equation}
%\label{eq:satelliteode}
%I\cdot \ddot{\theta} = \tau_{C}, 
%\end{equation}
%
%\noindent where $I$ is the satellite inertia moment about its mass
%center, $\tau_{C}$ is the control torque applied by the thrusters, and
%$\theta$ is the controlled attitude angle.  Normalizing
%Eq.~\eqref{eq:satelliteode} by defining the control signal
%$u=\frac{\tau_{C}}{I}$ and taking the Laplace transform, the following
%$s$-domain transfer function $G(s)$ is obtained:
%
%\begin{equation}
%\label{eq:satellitetf}
%G_{3}(s)=\frac{\theta(s)}{u(s)}=\frac{1}{s^2},
%\end{equation}
%
%\noindent Using the ZOH discretization defined in Eq.~\eqref{eq:pulsetf}, 
%we get the following $z$-domain model:
%
%\begin{equation}
%G_{3}(z)= \frac{T^{2}}{2} \frac{z+1}{(z+1)^{2}}. \nonumber
%\end{equation}
%
%
%\begin{equation}
%\label{exampleA}
%G_{4}(s)=\frac{s-1}{s^{2}-s-2},
%\end{equation}
%
%\noindent We can obtain $G_{4}(z)$ using the ZOH discretization defined
%in Eq.~\eqref{eq:pulsetf} with three different sample times ($0.1$, $0.05$,
%and $0.03$ seconds), respectively, as follows:
%
%\begin{equation}
%\label{exampleA-sampletime1}
%G_{4a}(z)=\frac{0.100342181002722z - 0.110876810062963}{1.0z^{2} - 2.12624017619613z + 1.10517091807565}, \nonumber
%\end{equation}
%
%
%\begin{equation}
%\label{exampleA-sampletime2}
%G_{4b}(z)=\frac{0.0500422033454653z - 0.0526068264456340}{1.0^{2} - 2.05640034257636z + 1.05127109637602}, \nonumber
%\end{equation}
%
%\begin{equation}
%\label{exampleA-sampletime3}
%G_{4c}(z)=\frac{0.0300090687252212z - 0.0309228417953966}{1.0^{2} - 2.03228208009387z + 1.03045453395352}. \nonumber
%\end{equation}

The fifth and sixth set of benchmarks describe the discrete model 
of a DC servo motor velocity dynamics~\cite{exampleCAD,Tan01}.

The seventh set of benchmarks contains the discrete well-studied 
non-minimal phase model $G_{7}(z)$. Non-minimal phase model offers 
additional difficulties for the controller design~\cite{Doyle:1991:FCT:574259}.

%The fifth set of benchmarks contains the \textit{DC Motor Rate} discrete model $G_{5}(z)$, which 
%describes the angular velocity of a DC Motor plant~\cite{Franklin15}. 
%
%\begin{equation}
%\label{DC-Motor-Rate}
%G_{5}(z)=\frac{ 0.1898z + 1.8027\times{10^{-4}}}{1.0z^{2} -0.9012z -1.0006\times{10^{-16}}}. \nonumber
%\end{equation}
%
The eighth set of benchmarks contains the \textit{Helicopter Longitudinal Motion} discrete model $G_{8}(z)$,
which provides the longitudinal motion dynamics of a helicopter~\cite{Franklin15}. 
%
%\begin{equation}
%\label{Helicopter-Longitudinal-Motion}
%G_{6}(z)=\frac{15.1315z^{2} + 17.8600z + 17.4549}{z^{3} -2.6207z^{2} + 2.3586z -0.6570}. \nonumber
%\end{equation}
%
The ninth  set of benchmarks contains the \textit{Inverted Pendulum} discrete model $G_{9}(z)$, which 
describes a pendulum dynamics with its center of mass above its pivot point~\cite{Franklin15}. 
%
%\begin{equation}
%\label{Inverted-Pendulum}
%G_{7}(z)=\frac{0.0263z^{2} + 0.0143z -0.0383}{z^{4} -4.2553z^{3} + 6.5105z^{2} -4.2553z + 1}. \nonumber
%\end{equation}
%
The tenth set of benchmarks contains the \textit{Magnetic Suspension} discrete model $G_{10}(z)$, which describes the dynamics of a mass that levitates with support only of a magnetic field~\cite{Franklin15}. 
%
%\begin{equation}
%\label{Magnetic-Suspension}
%G_{8}(z)=\frac{6.769\times{10^{13}}z + 6.784\times{10^{13}}}{z^{2} -5.415\times{10^{13}}z + 4.23\times{10^{11}}}. \nonumber
%\end{equation}

@@

We give the runtimes required to synthesize a stable controller for each
benchmark in Table~\ref{tab:results}.  Here, \textit{Plant} is the discrete
or continuous plant model, \textit{Benchmark} is the name of the employed
benchmark, \textit{I} and \textit{F} represent the number of integer and
fractional bits of the stable controller, respectively, while \textit{Time}
is the total time (in seconds) required to synthesize a stable controller
for the given plant.

For the majority of our benchmarks, the conjecture explained in
Section~\ref{synthesis-elements} holds and the two-stage verification 
engine %inductive back-end 
is able to find a stable solution in less than five minutes.  This is possible if the
inductive solutions need to be refined with few counterexamples and
increases of the fixed-point precision.  However, two benchmarks in our set
required too many counterexamples to refine their solution (marked with $^*$
in Table~\ref{tab:results}).  In these cases, the one-stage engine 
%interval arithmetic back-end 
is able to complement the two-stage approach and synthesize a solution
instead, albeit with significantly larger runtimes.  This performance
difference is due to the fact that the one-stage verification engine does
not take advantage of the inductive conjecture inherent to CEGIS, but
instead fully explores the counterexample space.  The result is a
performance difference of at least two orders of magnitude in our
experiments, leading to the one-stage engine timing out on the majority of
our benchmarks.  Table~\ref{tab:results} only lists the faster of the two
verification engines for each benchmark, which in $12$ out of $15$
benchmarks is the two-stage one.

@@

The median runtime for our benchmark set is $48$\,s, implying that \tool
can synthesize half of the controllers in less than one minute.  Overall,
the average synthesis time amounts to approximately $42$ minutes.  We
consider these times short enough to be of practical use to control
engineers, and thus affirm RQ1.  We further observe that the approach using
the two-stage verification engine is able to synthesize stable controllers
in $19$ out of the $23$ benchmarks, and can be complemented using the
one-stage engine for additional benchmarks, where the inductive
conjectures fail.  Both verification engines together enable controller
synthesis for $20$ out of $23$ benchmarks.  For the remaining benchmarks
our approach failed to synthesize a stable controller within the time
limits.  This can be addressed by either increasing either the time limit
or the fixed-point word widths considered, or by using floating-point
arithmetic instead.  The synthesized controllers were confirmed to be stable
outside of our model representation using MATLAB, positively answering RQ2.
A~link to the full experimental environment, including scripts to reproduce
the results, all benchmarks and the \tool tool, is provided in the
footnote.\footnote{\url{https://drive.google.com/file/d/0ByIexo3Z5N91QkhrMi1rdkZFTWc/view}}\footnote{CBMC version: \url{https://github.com/diffblue/cbmc/archive/7a6cec1dd0eb8843559591105235f1f2c4678801.zip}}

\subsection{Threats to validity}

>>>>

The first set of benchmarks uses the discrete model $G_1(z)$ of a cruise control
system for a car, and accounts for rolling friction, aerodynamic drag, and
the gravitational disturbance force~\cite{Astrom08}. 
%The plant model is given as follows:
%
%\begin{equation}
%\label{cruise-control-c1}
%G_1(z)=\frac{0.0264}{z-0.998} \nonumber
%\end{equation} 
%
The second set of benchmarks considers the discrete model $G_2(z)$ 
of a simple spring-mass damper plant~\cite{DBLP:conf/hybrid/WangGRJF16}. 
%where the discrete 
%plant dynamics is represented by the following $z$-expression:
%
%\begin{equation}
%\label{spring-mass-damper-g}
%G_2(z)=\frac{5\times{10^{-5}}z + 5\times{10^{-5}}}{z^2 - 2z + 1.0001}. \nonumber
%\end{equation} 
%
A third set of benchmarks uses the discrete model $G_{3}(z)$ for satellite attitude 
dynamics~\cite{Franklin15}, which require attitude (pose) control
for orientation of antennas and sensors w.r.t.~earth.
The fourth set of benchmarks presents an alternative discrete model $G_4(z)$ 
of a cruise control system~\cite{DBLP:conf/hybrid/WangGRJF16}. 
%The satellite
%attitude control is typically used for three-axis attitude tracking, but
%here we consider only one axis at a time.  
%The motion equation for the
%one-axis system, disregarding disturbance torques, is given by
%
%\begin{equation}
%\label{eq:satelliteode}
%I\cdot \ddot{\theta} = \tau_{C}, 
%\end{equation}
%
%\noindent where $I$ is the satellite inertia moment about its mass
%center, $\tau_{C}$ is the control torque applied by the thrusters, and
%$\theta$ is the controlled attitude angle.  Normalizing
%Eq.~\eqref{eq:satelliteode} by defining the control signal
%$u=\frac{\tau_{C}}{I}$ and taking the Laplace transform, the following
%$s$-domain transfer function $G(s)$ is obtained:
%
%\begin{equation}
%\label{eq:satellitetf}
%G_{3}(s)=\frac{\theta(s)}{u(s)}=\frac{1}{s^2},
%\end{equation}
%
%\noindent Using the ZOH discretization defined in Eq.~\eqref{eq:pulsetf}, 
%we get the following $z$-domain model:
%
%\begin{equation}
%G_{3}(z)= \frac{T^{2}}{2} \frac{z+1}{(z+1)^{2}}. \nonumber
%\end{equation}
%
%
%\begin{equation}
%\label{exampleA}
%G_{4}(s)=\frac{s-1}{s^{2}-s-2},
%\end{equation}
%
%\noindent We can obtain $G_{4}(z)$ using the ZOH discretization defined
%in Eq.~\eqref{eq:pulsetf} with three different sample times ($0.1$, $0.05$,
%and $0.03$ seconds), respectively, as follows:
%
%\begin{equation}
%\label{exampleA-sampletime1}
%G_{4a}(z)=\frac{0.100342181002722z - 0.110876810062963}{1.0z^{2} - 2.12624017619613z + 1.10517091807565}, \nonumber
%\end{equation}
%
%
%\begin{equation}
%\label{exampleA-sampletime2}
%G_{4b}(z)=\frac{0.0500422033454653z - 0.0526068264456340}{1.0^{2} - 2.05640034257636z + 1.05127109637602}, \nonumber
%\end{equation}
%
%\begin{equation}
%\label{exampleA-sampletime3}
%G_{4c}(z)=\frac{0.0300090687252212z - 0.0309228417953966}{1.0^{2} - 2.03228208009387z + 1.03045453395352}. \nonumber
%\end{equation}
%
The fifth and sixth set of benchmarks describe the discrete model 
of a DC servo motor velocity dynamics~\cite{exampleCAD,Tan01}.

The seventh set of benchmarks contains the discrete well-studied 
non-minimal phase model $G_{7}(z)$. Non-minimal phase model offers 
additional difficulties for the controller design~\cite{Doyle:1991:FCT:574259}.
%
%The fifth set of benchmarks contains the \textit{DC Motor Rate} discrete model $G_{5}(z)$, which 
%describes the angular velocity of a DC Motor plant~\cite{Franklin15}. 
%
%\begin{equation}
%\label{DC-Motor-Rate}
%G_{5}(z)=\frac{ 0.1898z + 1.8027\times{10^{-4}}}{1.0z^{2} -0.9012z -1.0006\times{10^{-16}}}. \nonumber
%\end{equation}
%
The eighth set of benchmarks contains the \textit{Helicopter Longitudinal Motion} discrete model $G_{8}(z)$,
which provides the longitudinal motion dynamics of a helicopter~\cite{Franklin15}. 
%
%\begin{equation}
%\label{Helicopter-Longitudinal-Motion}
%G_{6}(z)=\frac{15.1315z^{2} + 17.8600z + 17.4549}{z^{3} -2.6207z^{2} + 2.3586z -0.6570}. \nonumber
%\end{equation}
%
The ninth  set of benchmarks contains the \textit{Inverted Pendulum} discrete model $G_{9}(z)$, which 
describes a pendulum dynamics with its center of mass above its pivot point~\cite{Franklin15}. 
%
%\begin{equation}
%\label{Inverted-Pendulum}
%G_{7}(z)=\frac{0.0263z^{2} + 0.0143z -0.0383}{z^{4} -4.2553z^{3} + 6.5105z^{2} -4.2553z + 1}. \nonumber
%\end{equation}
%
The tenth set of benchmarks contains the \textit{Magnetic Suspension} discrete model $G_{10}(z)$, which describes the dynamics of a mass that levitates with support only of a magnetic field~\cite{Franklin15}. 
%
%\begin{equation}
%\label{Magnetic-Suspension}
%G_{8}(z)=\frac{6.769\times{10^{13}}z + 6.784\times{10^{13}}}{z^{2} -5.415\times{10^{13}}z + 4.23\times{10^{11}}}. \nonumber
%\end{equation}

@@

We give the run-times required to synthesize a stable controller for each
benchmark in Table~\ref{tab:results}.  Here, \textit{Plant} is the discrete
or continuous plant model, \textit{Benchmark} is the name of the employed
benchmark, \textit{I} and \textit{F} represent the number of integer and
fractional bits of the stable controller, respectively, while \textit{Time}
is the total time (in seconds) required to synthesize a stable controller
for the given plant.

For the majority of our benchmarks, the conjecture explained in
Section~\ref{synthesis-elements} holds and the two-stage verification 
engine %inductive back-end 
is able to find a stable solution in less than five minutes.  This is possible if the
inductive solutions need to be refined with few counterexamples and
increases of the fixed-point precision.  However, two benchmarks in our set
required too many counterexamples to refine their solution \textcolor{red}{
(marked with $^*$ in Table~\ref{tab:results})}.  
In these cases, the one-stage engine is able to complement the two-stage 
approach and synthesize a solution instead, albeit with significantly larger run-times.  
This performance difference is due to the fact that the one-stage verification engine does
not take advantage of the inductive conjecture inherent to CEGIS, but
instead fully explores the counterexample space.  The result is a
performance difference of at least two orders of magnitude in our
experiments, leading to the one-stage engine timing out on the majority of
our benchmarks. \textcolor{red}{Table~\ref{tab:results} only lists the faster of the two
verification engines for each benchmark, which in $12$ out of $15$
benchmarks is the two-stage one}.

@@

The median run-time for our benchmark set is $48$\,s, implying that \tool
can synthesize half of the controllers in less than one minute.  Overall,
the average synthesis time amounts to approximately $42$ minutes.  We
consider these times short enough to be of practical use to control
engineers, and thus affirm RQ1.  We further observe that the approach using
the two-stage verification engine is able to synthesize stable controllers
in $19$ out of the $23$ benchmarks, and can be complemented using the
one-stage engine for additional benchmarks, where the inductive
conjectures fail.  Both verification engines together enable controller
synthesis for $20$ out of $23$ benchmarks.  For the remaining benchmarks
our approach failed to synthesize a stable controller within the time
limits.  This can be addressed by either increasing either the time limit
or the fixed-point word widths considered, or by using floating-point
arithmetic instead.  The synthesized controllers were confirmed to be stable
outside of our model representation using MATLAB, positively answering RQ2.
A~link to the full experimental environment, including scripts to reproduce
the results, all benchmarks and the \tool tool, is provided in the
footnote.\footnote{\url{https://drive.google.com/file/d/0ByIexo3Z5N91QkhrMi1rdkZFTWc/view}}\footnote{CBMC version: \url{https://github.com/diffblue/cbmc/archive/7a6cec1dd0eb8843559591105235f1f2c4678801.zip}}

%--------------------------------------
\subsection{Threats to validity}
%--------------------------------------
